Accounting for uncertainty in simulation greenhouse gas emissions from dairy cattle in Kenya

### 2022-06-03 Manuscript_analyses.Rmd
cleaned up code - for comments

# Set up

Install and load required packages

@@

```{r ghg_emissions_function}
# Generate the GHG emissions function 
source("GHG_emissions_function.R", local = knitr::knit_global())
```

This equation takes a wide range of inputs. Many of these are usually not known precisely, yet the classic IPCC equation does not account for potential errors and uncertainties. Here, we express this uncertainty by defining distributions for each variable. The following table (**data/livestock_ghg_input_table.csv**) shows all the variables, as well as examples of distributions that may describe their quantities. Distributions are defined according to the distribution shape, as well as upper and lower bounds of 90% confidence intervals.

@@

The `scenario_mc` function in the `decisionSupport` package [@R-decisionSupport] consists of a wrapper around the `mcSimulation` function. The function can modify the input table according to a scenario file we provide. This file contains information on which variables should be modified for each scenario and how many model runs should be included in the simulation. The function is essentially the same as the `mcSimulation` function, but with an additional `scenarios` option which imports the contents of a .csv file.

The .csv file for the `scenarios` option specifies variables that should be modified for each of the scenarios, as well as the number of model runs for it. Each column beyond the first two columns in the file contains information for a particular scenario. This is restricted to the parameters that *should be changed* for this scenario.

@@

The first row ("Runs" in the first column) indicates the number of runs. Each subsequent row indicates which value from the input table should be modified, so the string in the first column has to correspond to a variable name in the input table. The second column indicates which values should be changed, with options being "lower", "upper", "distribution" and "both". If "both" is selected, both the lower and upper bounds are modified (this should only be done for variables witha constant distribution, since the rplacement will be for both the upper and lower bound).

We define subgroups based on the populations of households described by a household survey in Kenya [@wilkes_variation_2020; @wilkes_methods_2019]. Each of the farms encountered in this survey were converted into one scenario, for which Monte Carlo simulation is then used to produce multiple replicates.

@@

The information for the household is extracted from a spreadsheet that contains all the survey responses [@wilkes_variation_2020; @wilkes_methods_2019]. Extracting all relevant information from these spreadsheets is a slightly complex task, which is accomplished by the following code.

@@

We'll also make precise scenarios that contain only the exact values that were observed in the survey.

@@

@@

To illustrate the possible influence of input uncertainty on estimates of total farm emissions, we show the example of farm #5 from the household survey. This farm has 6 head of cattle, including 3 lactating cows. We simulate 1000 versions of this farm.

@@

The blue line represents a classic best-bet estimate based on default values in the IPCC guidelines and some context-specific reasonable guesses.

@@

We can illustrate this uncertainty using a scatter plot. The following figure shows all combinations of total milk yield and total greenhouse gas emissions across all households.

As with the GHG Model, we apply the uncertainty functions on the `milk_yield` and `on_farm` variables. These come from the model results (`r nrow(GHG_simulation_scenarios$y)`) of the `GHG_simulation_scenarios` model results to the `in_var` (x) and an outcome variable `out_var` (y). In the model we defined `on_farm` as the sum of Carbon Dioxide Equivalent Enteric Methane Emissions from Livestock (`enteric_CH4_CO2eq`), Carbon Dioxide Equivalent Methane Emissions from Livestock Manure (`mm_CH4_CO2eq`), Carbon Dioxide Equivalent Nitrous Oxide Emissions from Livestock Manure (`mm_N2O_CO2eq`) and Carbon Dioxide Emissions from Livestock Feed Production and Transport (`feed_CO2`).

```{r, warning=FALSE, message=FALSE}
plotdata<-data.frame(in_var=GHG_simulation_scenarios$y$pc_milk_yield,
                     out_var=GHG_simulation_scenarios$y$pc_on_farm,
                     Group="just one")

@@

An indicator that is often of interest is the ratio between these two variables: total emissions divided by the milk yield. We'll calculate this next. We then have to remove all households that don't produce any milk, because this ratio would be infinity for them. In fact, we'll remove all households that reported very low milk yields of <500 kg per year per capita, since these are unlikely to represent regularly operating dairy operations. For instance, they may be mostly focused on beef cattle production, or they may only have a small number of dairy cows that happened to be unproductive at the time of the survey.

@@

```{r varkernel}
#uncertainty::varkernel(in_var = dairy_households$y$pc_milk_yield,
#                        out_var = dairy_households$y$CO2eq_per_milk,
#                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
#                        ylab = expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]-eq~kg^{-1})))

@@

The milk yield per capita is sometimes proposed as an indicator of greenhouse gas emissions. We can now examine the merits of this indicator. Essentially, its usefulness hinges on its relationship with greenhouse gas emission intensity. We can investigate this by taking slices through the density surface.

@@

We see considerable uncertainty about the emissions at this particular level of production, and the previous figure shows that emissions of all other levels is similarly variable. This raises considerable doubts about the usefulness of milk yield as an indicator of emissions intensity - at least when this indicator is taken at face value, without considering this uncertainty. Yet, some information is contained even in such an imperfect indicator, since, as also shown in the previous figure, there is clearly a correlation between milk yield and emissions intensity. The following proposes a mechanism for quantifying the level of certainty, with which a change in milk yield can reasonably be associated with a change in emissions intensity.

@@

To derive the level of certainty, with which a change in milk yield is associated with an increase or decrease in emissions intensity, we have to investigate the probability distributions (as shown in Fig. X) of emissions intensity at two levels of milk production. This can be achieved using the `varslice_resample` function for both production levels.

@@

@@

The comparison in the last line of code above reveals that in `r sum(slice_3000$resampled < slice_2000$resampled) / length(slice_2000$resampled)*100` % of cases, emissions intensity at the higher milk production level is lower than at the lower level.

We can now generalize this result by doing similar calculations for multiple production levels and illustrating them for the whole population.

@@

decreased_emissions_probability[,"yield_increase"]<-decreased_emissions_probability$new_yield-decreased_emissions_probability$Base_yield

ggplot(data = decreased_emissions_probability, aes(Base_yield, yield_increase, z =
                                                      probability)) + geom_contour_filled() +
  ggtitle("Probability that emission intensity reductions occurred") + 
  xlab("Baseline milk yield per head and year (kg)") +
  ylab("Milk yield increase per head and year (kg)") +
  stat_contour(breaks = c(0.9),col="black")

ggsave("Fig_8_probability_of_intensity_reduction.png", width=7,height=5)

@@

The 90% confidence contour is shown in the plot. The results indicate that whenever milk yields increase, the probability that emissions intensity decreased is usually greater than 50% (slightly lower values are due to randomness in the resampling procedure), but that large increases are needed to be sure of this.

@@

Farms may try to reduce their emissions intensity by making changes to their herd structure, feeding strategy or management practices. Programs aiming to effect emissions reductions in dairy production may want to reward such changes, and they may be looking to the milk production per capita as an indicator for the effectiveness of such measures. We now want explore whether this is a reasonable strategy.

## Strategy

Develop scenarios that implement innovations

* breed adjustment
* herd structure

Quantify improvements on farms by comparing with the baseline.
Plot improvements vs. the baseline on the above figure.
Plot improvements vs. baseline in the emissions intensity figure.

## Single-run function

To clearly characterize the impact of interventions, we should compare their impact with a counterfactual, i.e. with the same farms without the intervention. While this is usually impossible in practice, it is easy with a model. We just have to run the simulation again with identical inputs except for those that are affected by the intervention.

We already simulated emissions for a large population of farms, and we can now apply our interventions to this population by modifying certain parameters of the input tables. What we don't want now, however, is to randomize all values again, since this would add additional noise to the results, which would mask the signal caused by the interventions.

We now first need a function that can apply our greenhouse gas emissions model to a pre-defined set of input values. We can find syntax that can achieve this in the mcSimulation function. Here it is (with slight modifications):

@@

The following code can precisely reproduce the results of the previous analysis:

@@

# Emission reductions through higher-performance breeds

Introduction of higher-performance breeds may reduce greenhouse gas emissions. We simulate an intervention that replaces animals of low-performance breeds with animals of high-performance breeds.

To retain the farm-specific performance levels obtained from the farm survey, we simulate these impacts in a relative manner. For this, we first need to determine the performance and characteristics of each of the breeds. We'll base this on the farm survey data (`Kenyaherds` dataset from above).

@@

ggsave("Fig_9_breeds.png", width=7,height=5)

milk_summary<-cbind(aggregate(cows$milk_yield, by=list(cows$breed), FUN=median),table(cows$breed))
milk_summary<-milk_summary[,c(1,4,2)]
colnames(milk_summary)<-c("Breed","n","Median_milk_yield")
milk_summary

@@

Let's define the intervention as follows:

All animals are replaced by high-performance animals of breed 1, which increases milk yields proportionally to the median milk yield differences. This means that milk yields are multiplied by the following factors:

@@

milk_summary             
```

@@

milk_change<-aggregate(Kenyaherds_interventions$milk_change[cows], by=list(Kenyaherds_interventions$QQNO[cows]), FUN=mean)

breed_intervention$milk_yield_4[which(breed_intervention$Farm %in% milk_change[,1])]<-milk_change[,2]
breed_intervention$milk_yield_5<-breed_intervention$milk_yield_4
breed_intervention$milk_yield_6<-breed_intervention$milk_yield_4

@@

Go through the x file, for each row take a random draw based on adoption rate (does farm adopt or not), then (possibly) apply factors. Actually, in this case we want to simulate full adoption, so we'll set this to 1.

@@

## Emissions intensity response to introduction of high-performance breeds

>>>>

# Setting up

Note that the code for all operations is hidden, so that it doesn't clutter the document. You can make it appear by clicking the code button on the right.

We first need to install and load all required packages.

@@

The equations provided in these documents, in particular in the 2006 guideline document [@dong_ipcc_2006] were converted into an R function called `ghg_emissions`, which is defined here (hit the `Code` button on the right).

```{r ghg_emissions_function}
# Generate the GHG emissions function 
ghg_emissions<-function(x, varnames){

  # Enteric fermentation emissions ($CH_4$)

  # The main source of all this is Chapter 10 of the *2006 IPCC Guidelines 
  # for National Greenhouse Gas Inventories* 
  # (Volume 4 - Agriculture, Forestry and Other Land Use). 
  # The document can be found on the IPCC web 
  # https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf

  # We start with the net energy for maintenance (NE_m; in MJ/day)
  # This includes a maintenance coefficient, 
  # for which different values are recommended depending on animal type

  # The following animal types were used in the example:

  # animal type 1 - 'intact' grown-up males >=36 months
  # animal type 2 - castrated grown-up males >=36 months
  # animal type 3 - males 12-36 months
  # animal type 4 - lactating cows
  # animal type 5 - lactating cows
  # animal type 6 - lactating cows
  # animal type 7 - weaned young females <= 12 months
  # animal type 8 - weaned young males <= 12 months
  # animal type 9 - females 12-36 months
  # animal type 10 - pre-weaned female calves (<12 months)
  # animal type 11 - pre-weaned male calves (<12 months)
  # animal type 12 - first-time pregnant cows

  # animal types 4, 5 and 6 are all lactating cows - 
  # important to note here that the original data table 
  # appears to have worked with the actual number of months 
  # that cows were pregnant. This resulted in a weighted mean 
  # of the various maintenance coefficients. 
  # We're not sure where we could get such data. 
  # In the original dataset, types 4, 5 and 6 
  # referred to different pregnancy/lactation states, 
  # but it doesn't seem like this scheme was rigorously followed.

  # We make placeholder vectors to collect emissions data 
  # for all animal types
  mm_N2O_CO2eq<-c()
  mm_CH4_CO2eq<-c()
  enteric_CH4_CO2eq<-c()
  total_milk<-c()

  # Now we start a long loop, 
  # where emissions are calculated for each animal type

  for (animal_type in 1:12)
  {
    # define C and Cfi based on sex (Page 10.17, below Equation 10.6)
    if(animal_type %in% c(1,3,8,11)) growth_coefficient_C<-C_intact_males # males (1.2)
    if(animal_type %in% c(2)) growth_coefficient_C<-C_castrates # castrates (1.0)
    if(animal_type %in% c(4,5,6,7,9,10,12)) growth_coefficient_C<-C_females # females (0.8)

    if(animal_type %in% c(4,5,6)) Cfi<-maintenance_coefficient_Cfi_lact_cow # lactating cows (0.386)
    if(animal_type %in% c(1)) Cfi<-maintenance_coefficient_Cfi_bulls # bulls (0.37)
    if(animal_type %in% c(2,3,7,8,9,10,11,12)) Cfi<-maintenance_coefficient_Cfi_other # others (0.322)

    # define mature body weight (assuming this needs to be disaggregated by sex, for Eqation 10.6)
    if(animal_type %in% c(1,2,3,8,11)) mature_weight_MW<-mature_weight_MW_male
    if(animal_type %in% c(4,5,6,7,9,10,12)) mature_weight_MW<-mature_weight_MW_female

    # define milk yield
    if(animal_type %in% c(1,2,3,7,8,9,10,11,12)) milk_yield<-0
    if(animal_type %in% c(4,5,6)) milk_yield<-eval(parse(text=paste0("milk_yield_",animal_type)))

    #define pregnancy coefficient (full Cp for first-time pregnant, lower for later in most cases)
    if(animal_type==12) Cp<-Cp_0
    if(animal_type==4) Cp<-Cp_0*Cp_rate_4
    if(animal_type==5) Cp<-Cp_0*Cp_rate_5
    if(animal_type==6) Cp<-Cp_0*Cp_rate_6
    if(animal_type %in% c(1,2,3,7,8,9,10,11)) Cp<-0

    N_animal_type<-eval(parse(text=paste0("N_animal_type_",animal_type)))

    # define activity coefficient Ca (Table 10.5)
    # Stall 0.00 - Animals are confined to a small area 
    #     (i.e., tethered, pen, barn) with the result that they expend 
    #     very little or no energy to acquire feed.
    # Pasture 0.17 - Animals are confined in areas with sufficient forage 
    #     requiring modest energy expense to acquire feed.
    # Grazing large areas 0.36 - Animals graze in open range land 
    #     or hilly terrain and expend significant energy to acquire feed.

    if(feeding_system==1) C_activity<-Ca_stall # Stall (0.0)
    if(feeding_system==2) C_activity<-Ca_pasture # Stall (0.17)
    if(feeding_system==3) C_activity<-Ca_large_grazing # Stall (0.36)

    # Mean live weight (kg) of animals surely varies by animal type, 
    # so we need separate estimates for them
    live_weight_LW<-eval(parse(text=paste0("live_weight_LW_",animal_type)))

    # Mean weight gain of animal type (kg/day)
    weight_gain_WG<-eval(parse(text=paste0("weight_gain_WG_",animal_type)))

    # Calculate the Net Energy for Maintenance (NE_m) (MJ/day) (Equation 10.3)
    NE_m<-Cfi*(live_weight_LW^0.75)

    # Calculate the Net Energy for Activity (NE_activity) (MJ/day) (Equation 10.4)
    NE_activity<-C_activity*NE_m

    #Net energy for growth (NE_growth) (MJ/day) (Equation 10.6)
    NE_growth<-22.02*
      ((live_weight_LW/(growth_coefficient_C*mature_weight_MW))^0.75)*
      (weight_gain_WG^1.097)

    # Net energy for lactation
    NE_lactation<-milk_yield*(1.47+0.40*milk_fat)

    #Energy need during pregnancy (Equation 10.13)
    NE_pregnancy<-Cp*NE_m

    # Ratio of net energy available in a diet for maintenance 
    # to digestible energy consumed (Equation 10.14)
    REM<-1.123-(4.092/1000*DE_perc)+(1.126/100000*DE_perc^2)-(25.4/DE_perc)

    # Ratio of net energy available for growth in a diet 
    # to digestible energy consumed (Equation 10.15)
    REG<-1.164-(5.160/1000*DE_perc)+(1.308/100000*DE_perc^2)-(37.4/DE_perc)

    # Gross Energy (GE) MJ/day
    NE_work<-0 # work animals aren't covered here, so we set this to zero
    GE<-(((NE_m+NE_activity+NE_lactation+NE_work+NE_pregnancy)/REM)+
           (NE_growth/REG))/(DE_perc/100)

    # CH4 emission factors for enteric fermentation from a livestock category 
    # (Equation 10.21)
    enteric_CH4<-(GE*Y_m*365)/55.65 # kg CH4 per animal per year

    # Compute the CO2-equivalent of methane emissions for enteric methane
    # This converts methane emissions to their Global Warming Potential 
    # (CO2-equivalent of CH4 is 25)
    enteric_CH4_CO2eq[animal_type]<-N_animal_type*enteric_CH4*25

    # Manure management methane ($CH_4$) and nitrous oxide ($N_2O$)
    # daily volatile solid excreted (Equation 10.24)

    volsol<-(GE*
               (1-(DE_perc/100))+
               (urinary_energy*GE))*
      ((1-ash_content)/18.45)

    # The maximum methane emission from manure (Bo) depends on the animal type. 
    # In the input table, we treat these as constants for each animal type. 
    # The following line assigns the corresponding Bo value

    Bo<-eval(parse(text=paste0("Bo_animaltype_",animal_type)))

    # CH4 emission factor from manure management (Equation 10.23)
    mm_ch4<-(volsol*365)*
      (Bo*0.67*
         ((mms_pasture*MCFpasture)+
            (mms_spread*MCFspread)+
            (mms_drylot*MCFdrylot)+
            (mms_solidstore*MCFsolidstore)+
            (mms_composted*MCFcomposted)+
            (mms_liquid*MCFliquid)+
            (mms_biogas*MCFbiogas)+
            (mms_burn*MCFburn)+
          # the original code was lacking the 'sold' fraction - 
          # we added this for completeness
          (mms_sold*MCFsold)))

    # Estimate CO2eq for manure methane
    mm_CH4_CO2eq[animal_type]<-N_animal_type*mm_ch4*25

    # Calculate Nitrous oxide emissions from manure and urine on pasture

    # estimate N intake (kg/animal/day) (Equation 10.32)
    N_intake<-(GE/18.45)*((perc_crude_protein_diet/100)/6.25)

    # estimate N retention (kg/animal/day) (Equation 10.33)

    if(weight_gain_WG==0)
      N_retention <- ((milk_yield*(1.9+0.4*milk_fat)/100)/6.38)

    if(!weight_gain_WG==0)
      N_retention<-((milk_yield*(1.9+0.4*milk_fat)/100)/6.38)+
      ((weight_gain_WG*(268-(7.03*NE_growth/weight_gain_WG)))/(1000*6.25))

    # estimate N excretion 
    # (kg N/animal/yr) (Equation 10.31)
    N_excretion<-N_intake*(1-N_retention)*365

    # Direct nitrous oxide emissions 
    # (kg N2O/animal/yr) (Eq 10.25 from IPCC guidelines)
    mm_direct_n2o<-N_excretion*((mms_pasture*mndec_pasture)+
                                  (mms_spread*mndec_spread)+
                                  (mms_drylot*mndec_drylot)+
                                  (mms_solidstore*mndec_solidstore)+
                                  (mms_composted*mndec_composted)+
                                  (mms_liquid*mndec_liquid)+
                                  (mms_biogas*mndec_biogas)+
                                  (mms_burn*mndec_burn)+
                                  (mms_sold*mndec_sold))*(44/28)

    # N losses from volatilization 
    # (kg N/animal/yr) (Eq 10.26 from IPCC guidelines)
    mm_vol_n<-N_excretion*((mms_pasture*mnvc_pasture)+
                             (mms_spread*mnvc_spread)+
                             (mms_drylot*mnvc_drylot)+
                             (mms_solidstore*mnvc_solidstore)+
                             (mms_composted*mnvc_composted)+
                             (mms_liquid*mnvc_liquid)+
                             (mms_biogas*mnvc_biogas)+
                             (mms_burn*mnvc_burn)+
                             (mms_sold*mnvc_sold))

    # indirect N2O emissions due to volatilization of N from manure management 
    # (kg N2O/animal/year) (Equation 10.27)
    mm_vol_n2o<-mm_vol_n*EF4*(44/28)

    # Nitrous oxide (N2O) from leaching (kg N/animal/year) (equation 10.28)
    mm_leach_n<-N_excretion*((mms_pasture*frac_leach_pasture)+
                               (mms_spread*frac_leach_spread)+
                               (mms_drylot*frac_leach_drylot)+
                               (mms_solidstore*frac_leach_solidstore)+
                               (mms_composted*frac_leach_composted)+
                               (mms_liquid*frac_leach_liquid)+
                               (mms_biogas*frac_leach_biogas)+
                               (mms_burn*frac_leach_burn)+
                               (mms_sold*frac_leach_sold))

    # indirect N2O emissions due to leaching of N from manure management 
    # (kg N2O/animal/year) (Equation 10.29)
    mm_leach_n2o<-mm_leach_n*EF5*(44/28)

    # CW EL Comment ####
    # # comment: we removed this, it was probably wrong, 
    # since leached nitrogen doesn't emit (much) nitrous oxide
    #mm_leach_n2o<-n_excretion*((mms_pasture*0.3)+(mms_spread*0.3)+(mms_drylot*0.3)+(mms_solidstore*0.3)+
    #              (mms_composted*0.3)+(mms_liquid*0.3)+(mms_biogas*0)+(mms_burn*0))*0.0075*(44/28)

    # Estimate CO2eq for manure nitrous
    mm_N2O_CO2eq[animal_type]<-N_animal_type*(mm_direct_n2o+mm_vol_n2o+mm_leach_n2o)*298 # (298 is the CO2-equivalent of N2O)

    total_milk[animal_type]<-N_animal_type*milk_yield*365

  # Emissions from feed production and transport.

  feed_CO2<-feed_prod_CO2+feed_trans_CO2

  # Total emissions

  on_farm<-sum(enteric_CH4_CO2eq+mm_CH4_CO2eq+mm_N2O_CO2eq)+feed_CO2

  total_animals<-sum(sapply(1:12,function(x) eval(parse(text=paste0("N_animal_type_",x)))))

  return(list(enteric_CH4=enteric_CH4,
              milk_yield=sum(total_milk),
              enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq),
              mm_CH4_CO2eq=sum(mm_CH4_CO2eq),
              mm_N2O_CO2eq=sum(mm_N2O_CO2eq),
              feed_CO2=feed_CO2,
              on_farm=on_farm,
              pc_enteric_CH4=enteric_CH4/total_animals,
              pc_milk_yield=sum(total_milk)/total_animals,
              pc_enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq)/total_animals,
              pc_mm_CH4_CO2eq=sum(mm_CH4_CO2eq)/total_animals,
              pc_mm_N2O_CO2eq=sum(mm_N2O_CO2eq)/total_animals,
              pc_feed_CO2=feed_CO2/total_animals,
              pc_on_farm=on_farm/total_animals

              ))
}

This equation takes a wide range of inputs. Many of these are usually not known precisely, yet the classic IPCC equation does not account for potential errors and uncertainties. Here, we express this uncertainty by defining distributions for each variable. The following table (**data/livestock_ghg_input_table.csv**) shows all the variables, as well as examples of distributions that may describe their quantities. Distributions are defined according to the distribution shape, as well as upper and lower bounds of 90% confidence intervals. We use the following distribution shapes:

* const - constants (upper and lower bounds are the same)
* posnorm - positive normal distribution; this is a normal distribution that is truncated at 0.

The decisionSupport package [@R-decisionSupport], which we use to run the simulations, also accommodates other distribution types, but we did not consider these appropriate here. The positive normal distribution (and other distributions such as normal or lognormal distributions) are defined based on the 5% and 95% quantiles. Note that it isn't always possible to fit a positive normal distribution based on such specifications, especially if the 5% quantile is set so close to zero that the remaining 5% of the data do not **fit** in the space between zero and the lower bound. In such cases, decisionSupport [@R-decisionSupport] returns a warning message, but it still does its best to generate the requested distribution. It is worth being on the lookout for the respective warning messages and try to adjust the bounds of the distribution to **possible** values.

@@

The `scenario_mc` function in the `decisionSupport` package [@R-decisionSupport] consists of a wrapper around the `mcSimulation` function. The function can modify the input table according to a scenario file we provide. This file contains information on which variables should be modified for each scenario and how many model runs should be included in the simulation. The function call is essentially the same as for the `mcSimulation` function, but with an additional `scenarios` option which imports the contents of a .csv file.

The .csv file for the `scenarios` option specifies variables that should be modified for each of the scenarios, as well as the number of model runs for it. Each column beyond the first two columns in the file contains information for a particular scenario. This is restricted to the parameters that *should be changed* for this scenario (all others don't have to be included).

@@

The first row ("Runs" in the first column) indicates the number of runs. Each subsequent row indicates which value from the input table should be modified, so the string in the first column has to correspond to a variable name in the input table. The second column indicates which values should be changed, with options being "lower", "upper", "distribution" and "both". If "both" is selected, both the lower and upper bounds are modified (this should only be done for variables with a constant distribution, since the replacement will be for both the upper and lower bound).

We define subgroups based on the populations of households described by a household survey in Kenya [@wilkes_variation_2020; @wilkes_methods_2019]. Each of the farms encountered in this survey was converted into one scenario, for which Monte Carlo simulation is then used to produce multiple replicates.

@@

The information for the household is extracted from a spreadsheet that contains all the survey responses [@wilkes_variation_2020; @wilkes_methods_2019].

### Probabilistic scenarios

Extracting all relevant information from the survey results spreadsheet is a slightly complex task, which is accomplished by the following code. The code includes specification of lower and upper bounds of probability distributions for most of the variables. We keep the herd composition constant, but all variables that in all likelihood weren't determined with high precision and accuracy are equipped with such uncertainty estimates. Estimates of the uncertainty for model input variables were extracted from the IPCC documents, or, if no such indications were available, based on our expert judgement.

@@

### Precise scenarios

We also make precise scenarios that contain only the exact values that were observed in the survey. This is needed for illustrating the impact of accounting for uncertainty later. For these scenarios, we follow the commonly used practice of applying default values recommended by the various IPCC documents mentioned above [@dong_ipcc_2006;@ipcc_2019;@ipcc_good_2021].

@@

## Running the simulation

@@

To illustrate the possible influence of input uncertainty on estimates of total farm emissions, we show the example of farm #5 from the household survey. This farm has 6 head of cattle, including 3 lactating cows. We simulate 1000 versions of this farm. We also do a precise calculation based on the exact numbers returned by the household survey and the default IPCC values.

### Total emissions of the farm

The following code and the resulting distribution estimate the total emissions of farm #5.

@@

The light blue distribution illustrates the estimated emissions from farm #5, considering uncertainty about all input variables. The blue line represents a classic best-bet estimate based on default values in the IPCC guidelines and some context-specific reasonable guesses.

### Emissions intensity

Naturally, the total emissions of the farm depend strongly on the size of the farm. A better indicator of the carbon footprint of milk production is the emissions intensity, i.e. the emissions per quantity of milk produced.

@@

We can illustrate this uncertainty using a scatter plot. The following figure shows all combinations of total milk yield and total greenhouse gas emissions across all households. Total emissions are the sum of methane emissions from enteric fermentation and livestock manure, nitrous oxide emissions from manure and carbon dioxide emissions from livestock feed production and transport.

```{r, warning=FALSE, message=FALSE}
plotdata<-data.frame(in_var=GHG_simulation_scenarios$y$pc_milk_yield,
                     out_var=GHG_simulation_scenarios$y$pc_on_farm)

@@

As mentioned above, an indicator that is often of interest is the ratio between total emissions and the milk yield. We'll calculate this next. We then have to remove all households that don't produce any milk, because this ratio would be infinity for them. In fact, we'll remove all households that reported very low milk yields of <500 kg per year per capita, since these are unlikely to represent regularly operating dairy operations. For instance, they may be mostly focused on beef cattle production, or they may only have a small number of dairy cows that happened to be unproductive at the time of the survey.

@@

```{r varkernel}

@@

The milk yield per capita is sometimes proposed as an indicator of greenhouse gas emissions. We can now examine the merits of this indicator. Essentially, its usefulness hinges on its relationship with greenhouse gas emissions intensity. We can investigate this by taking slices through the density surface.

@@

We see considerable uncertainty about the emissions at this particular level of production, and the previous figure shows that emissions of all other levels are similarly variable. This raises considerable doubts about the usefulness of milk yield as an indicator of emissions intensity - at least when this indicator is taken at face value, without considering this uncertainty. Yet, some information is contained even in such an imperfect indicator, since, as also shown in the previous figure, there is clearly a correlation between milk yield and emissions intensity. The following proposes a mechanism for quantifying the level of certainty, with which a change in milk yield can reasonably be associated with a change in emissions intensity.

@@

To derive the level of certainty, with which a change in milk yield is associated with an increase or decrease in emissions intensity, we have to investigate the probability distributions (as shown in the previous figure) of emissions intensity at two levels of milk production. This can be achieved by using the `varslice_resample` function for both production levels.

@@

slices<-rbind(data.frame(Values=slice_2000$resampled, Level="2000 kg"), data.frame(Values=slice_3000$resampled, Level="3000 kg"))

ggplot(data=slices, aes(Values, fill=Level)) +
 geom_density(alpha=.2) +
  ylab("Relative probability") +
    xlab(expression(Emission~intensity~of~milk~production~at~specific~milk~yield~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + labs(fill='Production level')

ggsave("Fig_8_emission_density_all_farms.png", width=7,height=5)

@@

Now we want to determine the probability of emissions intensity at a production level of 3000 kg milk per capita and year being lower than emissions intensity at 2000 kg per capita and year. We can do this empirically by comparing the two sets of 1000 random values that are assembled in the probability distribution slices representing per capita milk production levels of 3000 and 2000 kg. We test for each index number (1-1000) if the respective element of the 3000-kg slice is smaller than the respective element of the 2000-kg slice. By adding up the results of all these comparisons (with TRUE coded as 1 and FALSE as 0) and dividing by 1000 (and optionally multiplying by 100% for a percentage), we obtain the probability that emissions intensity is really lower at the higher milk production level.

In this case, this comparison reveals that in `r sum(slice_3000$resampled < slice_2000$resampled) / length(slice_2000$resampled)*100` % of cases, emissions intensity at the higher milk production level is lower than at the lower level.

We can now generalize this result by doing similar calculations for multiple production levels and illustrating them for the whole population. We'll base this on slices representing production levels between 100 and 10,000 kg per capita and year, at 100-kg intervals.

@@

#decreased_emissions_probability[,"yield_increase"]<-decreased_emissions_probability$new_yield-decreased_emissions_probability$Base_yield

ggplot(data = decreased_emissions_probability, aes(Base_yield, new_yield, z =
                                                      probability)) + geom_contour_filled() +
  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
       fill="Probability of \nlower emissions \nintensity") +
  stat_contour(breaks = c(0.9),col="black")

ggsave("Fig_9_probability_of_lower_emissions_intensity.png", width=7,height=5)

@@

The results indicate that whenever milk yields are higher on a particular farm than on another one, the probability that emissions intensity is lower is usually greater than 50% (slightly lower values are due to randomness in the resampling procedure), but that large differences are needed to be sure of this. The 90% confidence contour (shown in the plot as a black line) indicates the minimum level of milk yield that would be needed to state with 90% certainty that the emissions intensity is lower on the higher-productivity farm than on the respective lower-productivity one.

@@

Farms may try to reduce their emissions intensity by making changes to their herd structure, feeding strategy or management practices. Programs aiming to effect emissions reductions in dairy production may want to reward such changes, and they may be looking to the milk production per capita as an indicator for the effectiveness of such measures. We now want to explore whether this is a reasonable strategy.

## Interventions

We will explore the impacts of three interventions that farmers may consider. We'll restrict this assessment to interventions we can easily implement in our modeling framework.

### Breed adjustment

We evaluate milk production for all the breeds found in the survey dataset. All cows of low-productivity breeds are then replaced by high-productivity cows.

### Herd structure intervention 1 - retiring unproductive bulls

All bulls that are beyond their productive life span are removed from the herd. We implement this by removing all animals of animal type 1, which denotes bulls >36 months of age.

### Herd structure intervention 2 - fewer unproductive replacement males

Many farms have more male calves and young bulls than are needed to meet the objectives of dairy production. Based on a real-life scenario, we target a reduction in this number by 23%. For this, we randomly select 23% of the farms and set the number of replacement males (animal type 3) to 0.

## Analysis strategy

In a model, we can simulate the implications of specific interventions by leaving all variables unchanged, except those that are modified by the intervention. To implement this using the decisionSupport package, we have to disable randomness for the variables we don't want to modify. For this, we can create a function (`run_model`) that simply applies our model to a set of values we specify. We can use this function to exactly reproduce the results of a model run. Note that this only works, because our model is perfectly deterministic. All variation in model outputs results from variation in inputs; the model itself doesn't include additional random processes.

We can now generate the outputs of a Monte Carlo model for our baseline situation (without applying interventions). The output object of the `mcSimulation` function doesn't only contain the output variables, however, but also a complete list of all random inputs for all model runs. As mentioned above, we could now apply the `run_model` function to these inputs (the `x` element of the `mcSimulation` result) to reproduce the outputs of the simulation. More importantly, we can now modify certain variables included in this `x` element to simulate interventions, and then run the model on these modified inputs to simulate the impacts of these interventions.

The following code contains the `run_model` function:

@@

The following code can therefore precisely reproduce the results of the previous analysis:

@@

# Upgrading to higher-performance breeds

Replacing cows of low-productivity breeds with better-performing ones may reduce greenhouse gas emissions. We simulate an intervention that replaces animals of low-performance breeds with animals of high-performance breeds.

We first need to determine the performance and characteristics of each of the breeds that are present in the farm survey dataset (`Kenyaherds` dataset from above).

@@

ggsave("Fig_10_breeds.png", width=7,height=5)

milk_summary<-cbind(aggregate(cows$milk_yield, by=list(cows$breed), FUN=median),table(cows$breed))
milk_summary<-milk_summary[,c(1,4,2)]
colnames(milk_summary)<-c("Breed","n","Median_milk_yield")

milk_summary  %>%
    kableExtra::kbl(digits = 2)  %>%
    kableExtra::kable_classic_2(full_width = F)

@@

We define the intervention as follows:

All animals are replaced by high-performance animals of breed 1, which increases milk yields proportionally to the median milk yield differences. This means that milk yields are divided by the following factors:

@@

milk_summary   %>%
    kableExtra::kbl(digits = 2)  %>%
    kableExtra::kable_classic_2(full_width = F)            
```

@@

milk_change<-aggregate(Kenyaherds_interventions$milk_change[cows], by=list(Kenyaherds_interventions$QQNO[cows], Kenyaherds_interventions$animal_type[cows]), FUN=mean)

Milk_changes_4<-milk_change[which(milk_change$Group.2==4),]
breed_intervention$milk_yield_4[which(breed_intervention$Farm %in% Milk_changes_4[,1])]<-Milk_changes_4[,3]
Milk_changes_5<-milk_change[which(milk_change$Group.2==5),]
breed_intervention$milk_yield_5[which(breed_intervention$Farm %in% Milk_changes_5[,1])]<-Milk_changes_5[,3]
Milk_changes_6<-milk_change[which(milk_change$Group.2==6),]
breed_intervention$milk_yield_6[which(breed_intervention$Farm %in% Milk_changes_6[,1])]<-Milk_changes_6[,3]

@@

We can now go through the `x` file, and for each row take a random draw based on an adoption rate we can specify (does the farm adopt the intervention or not), then (possibly) apply factors. In this case we want to simulate full adoption, so we'll set this to 1.

@@

## Emissions intensity response to upgrading breeds

Exchanging cows of unproductive breeds to more productive individuals should have an effect on emissions intensity, yet this effect is difficult to see at the population level, both in the scatter plot and in the surface plot:
